NMRlipids IV: Headgroup \& glycerol backbone structures, and cation binding in bilayers with PE and PG lipids

### 2020-05-25 Manuscript/manuscriptPGPE.tex
Add data from MacRog POPE

@@

      POPE             & LIPID17 \cite{??} & 0      & 500 & 25000 & 50  &  310  & 500 & 100 & \cite{POPElipid17} \\
      POPE             & LIPID17 \cite{??} & 0.11   & 500 & 25000 & 50  &  310  & 500 & 100 & \cite{POPElipid17150mMNaCl} \\

@@

  \todo{We need citations for the force fields.}

@@

@@

\begin{figure}[]
  \centering
  \includegraphics[width=9.0cm]{../Figs/HGorderparametersPE.pdf}
  \caption{\label{HGorderParametersPE}
    The headgroup and glycerol backbone order parameters of PE lipids
    from experiments (POPE and signs this work, DPPE from Ref.~\citenum{seelig76}
    and E.coliPE from Ref.~\citenum{gally81}) and simulations with different force fields.
  }
  \todo{This should be clarified as in NMRlipidsI and error bars should be added.
    Probably larger error bars for united atom models based on the report by Fuchs et al.
  }
\end{figure}

@@

\clearpage
\subsection{Sodium binding to PE and PG lipid bilayers}
Sodium binding affinity to PE lipids has not been measured experimentally,
but large differences to PC would be surprising. 
The sodium binding affinity to POPE depends on the used force field (Fig. \ref{NAdensPE}),
but lesser extend than reported previously for PC~\cite{catte16}.
%Similarly to POPG,
Because some simulation and ion parameters are not identical with the previous work~\cite{catte16}
\todo{This should be finished once we have the simulation details},
we compare POPE results to the POPC simulations ran with identical parameters (Fig. \ref{NAdensPC}).
In Lipid17 with the strongest sodium binding affinity to POPE,
the binding affinity is approximately similar to POPC.
Slipids and CHARMM36 exhibit slightly, and
GROMOS-CKP subtantially weaker binding to POPE than to POPC.
%The Lipid17 exhibits stronger sodium binding affinity
%to POPE bilayers than Slipids, CHARMM36 and GROMOS-CKP (Fig. \ref{NAdensPE}).
%Experimental data for sodium binding to PE lipids is not available, while
%simulations in agreement with electrometer data from NMR suggests that
%sodium binding to POPC is weaker than in any of the simulations
%here (Fig. \ref{NAdensPC})~\cite{catte16,melcr18}.
Assuming that the binding
to POPE would be similar than to POPC, the sodium binding affinity to POPE
is potentially realistic in CHARMM36, Slipids, and GROMOS-CKP simulations here,
but substantially overestimated in Lipid17 simulation.
%On the other hand, the sodium binding affinity is stronger to POPE in Lipid17 simulations than to POPC in simulations with
%previous Amber lipid version, Lipid14.
%\todo{The potential differences in the used ion models should be checked}.
\begin{figure}[]
  \centering
  \includegraphics[width=8.0cm]{../Figs/NAdensPE.eps}
  \caption{\label{NAdensPE}
    Sodium (solid line) and choride ion density profiles along membrane normal
    from different simulations with PE lipids.
  }
\end{figure}

Simulations with PG lipids give similar dependence on force field as observed in POPE simulations:
Lipid 17 simulations exhibits stronger counter-ion binding affinity to pure POPG bilayer
than CHARMM36, Slipids, and GROMOS-CKP simulations, which are roughly similar (Fig. \ref{CIdensPG}).
Lipid17 also exhibits less increase in POPC headgroup
%$\beta$ carbon
order parameters upon addition of POPG than other simulations (Fig. \ref{HGorderparametersPCvsPEPG}),
and lower area per molecule (59.5 {\AA}$^2$) than in experiments (66.1 {\AA}$^2$).
In our previous study about PS lipids \cite{antila18}, such behaviour was related to 
%the results suggest that
the overestimated counterions binding and shielding the electrostatic repulsion between PG headgroups
in bilayers.% simulated with Lipid17 parameters.
%the underestimated increase of PC
%headgroup order parameters upon addition of negaticely charged lipids
%was related
%to the overestimated counterion binding affinity, which would overcompensates the effect of negative charge.
%Similar correlation is observed here in %other than GROMOS-CKP simulation:
%%Lipid 17 simulations, which exhibits stronger counter-ion binding affinity to pure POPG bilayer (Fig. \ref{CIdensPG})
%
Even though the area per lipid in CHARMM36, Slipids, and GROMOS-CKP simulations is in good agreement
with experiments (Fig. \ref{CIdensPG}), the experimental increase in 
%In GROMOS-CKP simulations, however, the $\beta$-carbon order parameter of
POPC headgroup order parameters upon addition of POPG are not fully reproduced (Fig. \ref{HGorderparametersPCvsPEPG}).
%more than in CHARMM36 and Slipids simulations even though the binding affinities in these
%simulations are similar. On the other hand, the responses of different hydrogens attached in $\alpha$-carbons
%are different in all simulations except in Slipids, most sever in Lipid17 and GROMOS-CKP where other in increases
%and other decreases \todo{Can we figure out the reason for this?}.
Therefore we conclude that the counter-ion binding affinity is overestimated in Lipid17 simulations,
% with the strongest affinity among the tested models,
while the other simulations are more realistic, but slight overbinding cannot be excluded.
\begin{figure}[]
  \centering
  \includegraphics[width=9.0cm]{../Figs/CIdensPG.eps}
  \caption{\label{CIdensPG}
    Counterion densities and area per lipids from simulations with PG lipids.
    Experimental area for POPG at 303 K is 66.1 {\AA}$^2$ and 67 {\AA}$^2$ for DPPC at 323 K \cite{pan12b}.
  }
\end{figure}

>>>>

      POPE  & OPLS-MacRog \cite{rog16}     &0    & 128	& 5120     &0    & 300  & 500 & 300 & \cite{MacRogPOPEfilesT300K} \\

@@

      POPE             & LIPID17 \cite{gould18} & 0      & 500 & 25000 & 50  &  310  & 500 & 100 & \cite{POPElipid17} \\
      POPE             & LIPID17 \cite{gould18} & 0.11   & 500 & 25000 & 50  &  310  & 500 & 100 & \cite{POPElipid17150mMNaCl} \\

@@

  \todo{Citation for CHARMM36 PE?} \\
  \todo{Which ion model is used in \cite{POPEcharmm150mMNaCl}?} \\
  \todo{Citation for GROMOS-CKP?} \\
  \todo{Citation for GROMOS 43A1-S3?} \\
  \todo{Citation for OPLS-UA models?} \\
  \todo{Citations for Berger-* simulations?} \\
  \todo{LIPID17 simulations with correct dihedrals still coming}

@@

      POPC             & OPLS-MacRog \cite{rog16}     &0          & 0  & 128 & 5120 & 0  &  300  & 500 & 300 & \cite{POPCmacrog300K} \\
      POPC:POPE (1:1)  & OPLS-MacRog \cite{rog16}     &0          & 0  & 128 & 5120 & 0  &  300  & 500 & 300 & \cite{POPC1POPE1macrogT300K} \\
      POPC             & Slipid \cite{jambeck12b}     &0          & 0  & 512 & 23943 & 0  &  300  & 170 & 100 & \cite{POPCslipid300K} \\
      POPC:POPE (1:1)  & Slipid \cite{jambeck12b}     &0          & 0  & 128 & 5120 & 0  &  300  & 500 & 300 & \cite{POPC1POPE1slipidT298K} \\

@@

\begin{figure}[]
  \centering
  \includegraphics[width=9.0cm]{../Figs/HGorderparametersPE.pdf}
  \caption{\label{HGorderParametersPE}
    The headgroup and glycerol backbone order parameters of PE lipids
    from experiments (POPE and signs this work, DPPE from Ref.~\citenum{seelig76}
    and E.coliPE from Ref.~\citenum{gally81}) and simulations with different force fields.
  }
  \todo{This should be clarified as in NMRlipidsI and error bars should be added.
    Probably larger error bars for united atom models based on the report by Fuchs et al.
  } \\
\end{figure}

@@

\clearpage
\subsection{Sodium binding to PE and PG lipid bilayers}
Sodium binding affinity to PE lipids has not been measured experimentally,
but large differences to PC would be surprising. 
In simulations, the sodium binding affinity to POPE depends on the used force field (Fig. \ref{NAdensPE}),
but lesser extend than reported previously for PC~\cite{catte16}.
\todo{This will be finished once we have all the simulation details and Lipid17 simulations with correct dihedrals from issue https://github.com/NMRLipids/NMRlipidsIVPEandPG/issues/12},
Because some simulation and ion parameters are not identical with the previous work~\cite{catte16},
we compare POPE results to the POPC simulations ran with identical parameters (Fig. \ref{NAdensPC}).
In Lipid17 with the strongest sodium binding affinity to POPE,
the binding affinity is approximately similar to POPC.
Slipids and CHARMM36 exhibit slightly, and
GROMOS-CKP subtantially weaker binding to POPE than to POPC.
%The Lipid17 exhibits stronger sodium binding affinity
%to POPE bilayers than Slipids, CHARMM36 and GROMOS-CKP (Fig. \ref{NAdensPE}).
%Experimental data for sodium binding to PE lipids is not available, while
%simulations in agreement with electrometer data from NMR suggests that
%sodium binding to POPC is weaker than in any of the simulations
%here (Fig. \ref{NAdensPC})~\cite{catte16,melcr18}.
Assuming that the binding
to POPE would be similar than to POPC, the sodium binding affinity to POPE
is potentially realistic in CHARMM36, Slipids, and GROMOS-CKP simulations here,
but substantially overestimated in Lipid17 simulation.
%On the other hand, the sodium binding affinity is stronger to POPE in Lipid17 simulations than to POPC in simulations with
%previous Amber lipid version, Lipid14.
%\todo{The potential differences in the used ion models should be checked}.
\begin{figure}[]
  \centering
  \includegraphics[width=8.0cm]{../Figs/NAdensPE.eps}
  \caption{\label{NAdensPE}
    Sodium (solid line) and choride ion density profiles along membrane normal
    from different simulations with PE lipids.
  }
\end{figure}

Simulations with PG lipids give similar dependence on force field as observed in POPE simulations:
Lipid 17 simulations with Dang ion parameters exhibits stronger counter-ion binding affinity to pure POPG bilayer
than CHARMM36, Slipids, and GROMOS-CKP simulations, which are roughly similar (Fig. \ref{CIdensPG}).
Lipid17 also exhibits less increase in POPC headgroup
%$\beta$ carbon
order parameters upon addition of POPG than other simulations (Fig. \ref{HGorderparametersPCvsPEPG}),
and lower area per molecule (59.5 {\AA}$^2$) than in experiments (66.1 {\AA}$^2$).
In our previous study about PS lipids \cite{antila18}, such behaviour was related to 
%the results suggest that
the overestimated counterions binding and shielding the electrostatic repulsion between PG headgroups
in bilayers.% simulated with Lipid17 parameters.
%the underestimated increase of PC
%headgroup order parameters upon addition of negaticely charged lipids
%was related
%to the overestimated counterion binding affinity, which would overcompensates the effect of negative charge.
%Similar correlation is observed here in %other than GROMOS-CKP simulation:
%%Lipid 17 simulations, which exhibits stronger counter-ion binding affinity to pure POPG bilayer (Fig. \ref{CIdensPG})
%
Even though the area per lipid in CHARMM36, Slipids, and GROMOS-CKP simulations is in good agreement
with experiments (Fig. \ref{CIdensPG}), the experimental increase in 
%In GROMOS-CKP simulations, however, the $\beta$-carbon order parameter of
POPC headgroup order parameters upon addition of POPG are not fully reproduced (Fig. \ref{HGorderparametersPCvsPEPG}).
%more than in CHARMM36 and Slipids simulations even though the binding affinities in these
%simulations are similar. On the other hand, the responses of different hydrogens attached in $\alpha$-carbons
%are different in all simulations except in Slipids, most sever in Lipid17 and GROMOS-CKP where other in increases
%and other decreases \todo{Can we figure out the reason for this?}.
Therefore we conclude that the counter-ion binding affinity is overestimated in Lipid17 simulations,
% with the strongest affinity among the tested models,
while the other simulations are more realistic, but slight overbinding cannot be excluded.
\begin{figure}[]
  \centering
  \includegraphics[width=9.0cm]{../Figs/CIdensPG.eps}
  \caption{\label{CIdensPG}
    Counterion densities and area per lipids from simulations with PG lipids.
    Experimental area for POPG at 303 K is 66.1 {\AA}$^2$ and 67 {\AA}$^2$ for DPPC at 323 K \cite{pan12b}.
  }
\end{figure}
